
The results can be grouped according to their combined correctness R2, and fall into four groups (see \autoref{table:LLM correctness groups}). 
The first group (\texttt{A}) with 27 points were able to respond correctly to all three questions. 

\medskip 

A second group (\texttt{B}) with combined correctneess of 27, 23 and 19 responded to all three question, but some question were only
answered partially. A third group (\texttt{C}) with combined correctness of 18, 9 answered correctly one or two question, where the last question
was always answered correctly. A fourth group (\texttt{D}) with combined correctness of 5 or 0 was not able to answer any of the questions
fully correctly or at all.

\medskip 
A second way of ranking model performance is according to R1, the difference between symbolic rating and prose rating.
Within group \texttt{A}, there only a small subset of six models had a significantly higher symbolic rating than prose rating, meaning they responded precisely in symbolic notation
without prose ($R1 \geq 7$). The highest R1 rating were scored by nvidia/nemotron-3-nano-30b-a3b (reasoning) with $R1 = 9$, 
google/gemini-3-flash-preview, although these models did not provide a correct and complete answer. 

\medskip 

In model ranks where both reasoning and non-reasoning can be compared, the runs with reasoning mode enabled did not show
a large impact on correctness or rating than their non-reasoning counterparts with the following exceptions.
\begin{itemize}
\item \texttt{x-ai/grok-4.1-fast} reached a better ranking on symbolic output, but was in all cases correct and complete.
\item The resonaning mode of \texttt{deepseek/deepseek-v3.2} achieved a higher symbolic rating, but was only partially correct on one question
\item For \texttt{google/gemini-3-flash-preview} both reasoning and non-reasoning modes achieved the same symbolic rating, but reasoning mode only answered two of three questions.
\item \texttt{nvidia/nemotron-3-nano-30b-a3b} achieved a much higher symbolic rating, but only answered the last question in either modes.
\end{itemize}

Few models which produced correct output did not respond in UTF-8 regarding symbolic notation but in \LaTeX{} notation (Example: \texttt{deepseek/deepseek-v3.2})
One model in the set did not produce any meaningful output: \texttt{meta-llama/llama-3.2-1b-instruct}.
All models except \texttt{qwen/qwen3-max}, \texttt{meta-llama/llama-3.2-3b-instruct}, \texttt{meta-llama/llama-3.2-1b-instruct}, \texttt{google/gemma-3n-e4b-it} 
answered at least the last questions correctly, either in symbolic notation or in prose.
Although some models understand a reasoning value as input parameter, they do not work in reasoning mode and do not indicate an amount of reasoning tokens used.

The selection of models include both Open Source language models where the number of parameters are known, and closed frontier models where this number
is typically not disclosed. Thus we cannot analyse in detail how the amount of parameters influences the correcteness of the output, or the symbolic/prose ratings.
While many models with parameter sizes of e.g. 120b, 24b, 20b, 12b, 9b answer at least the last question correctly, and in several cases also partially on the first two questions,
the frontier models (with few exceptions) answer all three correctly. 

\begin{table}[h]
\centering
\caption{Model groups by completeness ($R2$), A: all answered correctly and completely, B: all answered, some of which only partially, C: some not answered D: none answered fully or correctly }
\label{tab:groups}
\begin{tabular}{cl}
\hline
\textbf{Group} & \textbf{Models} \\
\hline
\multirow{10}{*}{\textbf{A}} 
  & anthropic/claude-sonnet-4.5 \\
  & anthropic/claude-opus-4.5 \\
  & anthropic/claude-opus-4.5 (+R) \\
  & anthropic/claude-sonnet-4.5 (+R) \\
  & google/gemini-3-pro-preview (+R) \\
  & x-ai/grok-4.1-fast \\
  & x-ai/grok-4.1-fast (+R) \\
  & qwen/qwen3-coder \\
  & deepseek/deepseek-v3.2 \\
  & meta-llama/llama-4-maverick \\
\hline
\multirow{7}{*}{\textbf{B}}
  & google/gemini-3-flash-preview \\
  & deepseek/deepseek-v3.2 (+R) \\
  & meta-llama/llama-4-maverick (+R) \\
  & google/gemma-3n-e4b-it \\
  & mistralai/ministral-8b-2512 \\
  & mistralai/mistral-small-3.2-24b-instruct \\
  & google/gemini-3-flash-preview (+R) \\
\hline
\multirow{13}{*}{\textbf{C}}
  & nvidia/nemotron-3-nano-30b-a3b (+R) \\
  & openai/gpt-5.2 (+R) \\
  & openai/gpt-oss-20b (+R) \\
  & openai/gpt-5.2 \\
  & openai/gpt-oss-120b (+R) \\
  & anthropic/claude-haiku-4.5 (+R) \\
  & anthropic/claude-haiku-4.5 \\
  & nvidia/nemotron-3-nano-30b-a3b \\
  & mistralai/mistral-large-2512 \\
  & qwen/qwen3-32b \\
  & nvidia/nemotron-nano-9b-v2 (+R) \\
  & nvidia/nemotron-nano-12b-v2-vl (+R) \\
  & qwen/qwen3-max \\
\hline
\multirow{5}{*}{\textbf{D}}
  & meta-llama/llama-3.2-3b-instruct \\
  & google/gemma-3n-e4b-it \\
  & meta-llama/llama-3.2-1b-instruct \\
  & mistralai/ministral-8b-2512 \\
  & qwen/qwen3-coder \\
\hline
\end{tabular}
\label{table:LLM correctness groups}
\end{table}


