
Regarding the two research questions posed, it can be concluded that Large Language Models demonstrate varying degress
of correctness and ability to produce symbolic output across different model architectures and configurations. In particular, frontier models 
e.g. from Anthropic, Google and xAI show a strong performance in both correctness and symbolic output capabilities. 
Other models, e.g. from Qwen, Deepseek or Meta show a mixed performance of high correctness but lower ratings on symbolic notation capabilities.

\medskip

The analysis reveals that while some models excel in both correctness and symbolic output, others struggle with either aspect. 
The reasoning mode's effectiveness is model-dependent, and its impact on performance is not consistent across all models. 
The limited scope of the study and the variability in model performance make it difficult to draw 
definitive conclusions about the relationship between parameter count and model capabilities. 

\medskip

The effect of many LLMs answering only the last question (either correctly or at all) indicates that the request prompt should be improved
to allow language models to produce more comprehensive responses.

\medskip

This research suggest that for agentic processing of mathematical notation, frontier models of group \texttt{A} presented are the optimal choice, 
as they answer correctly and in a very ligh level of symbolic notation, leaving prose processing only in the internal reasoning tokens but rarely in the
output. This makes post-processing of LLM output more straightforward and efficient.